Scalable and Parameter-Free Exploration in Deep Reinforcement Learning

### 2022-01-06 papers/main.tex
TEMPORARY DEPLOY COMMIT

\title{Scalable and Parameter-Free Optimal Regret Reinforcement Learning}

@@

\section{Background}
A reinforcement learning environement is modelled as a Markov decision process (MDP)
\(M = \langle \state, \action, r, P, P_0, \gamma \rangle\), where \(\state\) is the
state space and \(\action\) is the set of available actions. At time \(t=0\) a state
\(s_0\) is sampled from the distribution \(P_0(\cdot)\). At each timestep an action
\(a_t \sim \pi(\cdot \vert s_t)\) is selected and the agent transitions to a new state
state \(s_{t+1} \sim P(\cdot \vert s_t, a_t)\). A scalar reward
\(r_{t+1} = r(\cdot \vert s_{t+1}, s_t, a_t)\) is observed. As the agent and environement
interact in a sequence of time steps, a history of observations
\(\mathcal{H}_t = (s_0, a_0, s_1, r_1, a_1, s_2, r_2, \dots, s_t, r_t)\) is collected.
The goal is to find a policy \(\pi^\star\), such that sampling actions
\(a \sim \pi^*(\cdot \vert s)\) maximizes the expected accumulated and discounted future reward,
\(J^\pi \defeq \E_\pi \left[ \sum_{t=1}^\infty \gamma^t r_t \right]\). An efficient
agent must be able to learn from the data it collects, but since the data is
dependent on the policy, it must also prioritize to explore states and actions that
the agent can learn a lot from.

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward for an optimal policy.
An learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is called Thompson sampling.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution, and selecting the action that gave the highest
reward. Bayesian methods been also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert s_0=S, a_0=A \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_{t+1}, s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. We can't find \(Q\) exactly, but with a good
approximation this policy, \(\pi(a \vert s) = \arg\max_a Q(s,a)\), should
hopefully be close to optimal. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value.
\citet{fortunato_noisy_2019} introduced NoisyNet for exploration.
These are networks with stochastic weights, where each weight \(w^{(l)}_{ij}\)
has an added pertrubaition sampled from a noise distribution with standard deviation \(\sigma^{(l)}_{ij}\).
After initializing \(\bm{w}\) and \(\bm{\sigma}\) such that the network has sufficient
stochasticity for exploration, both parameters are learned using standard backpropagation.
The approach is similar to variational inference schemes such as Bayes by backprop~\citep{blundell_weight_2015} where the weights of a neural network model are assumed to be normally distributed with mean 
\(\mu^{(l)}_{ij}\) and standard deviation \(\sigma^{(l)}_{ij}\). The objective of
Bayes by backprop, however, is to approximate the posterior distribution
\(p(\bm{w} \vert \mathcal{D})\) in a task with a fixed dataset \(\mathcal{D}\).
The parameter distribution in NoisyNet does not necessarily
converge to an approximate posterior. This means that it does not have the same
guarantee on total regret as methods that approximate a posterior over the
value functions~\citep{osband_generalization_2016}. They do, however, have
experimental results which shows that the function does not always converge to
a deterministic solution, but it is unclear why the network learns to introduce
more noise into the network parameters.

@@

@@

\begin{itemize}
    \item Optimal regret \textcolor{red}{TODO: define optimal regret}
    \item Incorporate domain knowledge
\end{itemize}

@@

\section{Experiments}
\subsection{Uninformative Prior}
\subsection{Informative Prior}

>>>>

\title{Scalable and Parameter-Free Exploration in Deep Reinforcement Learning}

@@

\section{Introduction}
A reinforcement learning environement is modelled as a Markov decision process (MDP)
\(M = \langle \state, \action, r, P, P_0, \gamma \rangle\), where \(\state\) is the
state space and \(\action\) is the set of available actions. At time \(t=0\) a state
\(s_0\) is sampled from the distribution \(P_0(\cdot)\). At each timestep an action
\(a_t \sim \pi(\cdot \vert s_t)\) is selected and the agent transitions to a new state
state \(s_{t+1} \sim P(\cdot \vert s_t, a_t)\). A scalar reward
\(r_t = r(\cdot \vert s_t, a_t)\) is observed. As the agent and environement
interact in a sequence of time steps, a history of observations
\(\mathcal{H}_t = (s_0, a_0, r_0, s_1, a_1, r_1, \dots, s_t, a_t, r_t)\) is collected.
The goal is to find a policy \(\pi^\star\), such that sampling actions
\(a \sim \pi^*(\cdot \vert s)\) maximizes the expected accumulated and discounted future reward,
\(J^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \right]\). An efficient
agent must be able to learn from the data it collects, but since the data is
dependent on the policy, it must also prioritize to explore states and actions that
the agent can learn a lot from.

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
An learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution, and selecting the action that gave the highest
reward. Bayesian methods been also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert s_0=S, a_0=A \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value function.

\citet{fortunato_noisy_2019} introduced NoisyNet for exploration.
These are networks with stochastic weights, where each weight \(w^{(l)}_{ij}\)
has an added pertrubaition sampled from a noise distribution with standard deviation \(\sigma^{(l)}_{ij}\).
After initializing \(\bm{w}\) and \(\bm{\sigma}\) such that the network has sufficient
stochasticity for exploration, both parameters are learned using standard backpropagation.
The approach is similar to variational inference schemes such as Bayes by backprop~\citep{blundell_weight_2015} where the weights of a neural network model are assumed to be normally distributed with mean 
\(\mu^{(l)}_{ij}\) and standard deviation \(\sigma^{(l)}_{ij}\). The objective of
Bayes by backprop, however, is to approximate the posterior distribution
\(p(\bm{w} \vert \mathcal{D})\) in a task with a fixed dataset \(\mathcal{D}\).
The parameter distribution in NoisyNet does not necessarily
converge to an approximate posterior. This means that it does not have the same
guarantee on total regret as methods that approximate a posterior over the
value functions~\citep{osband_generalization_2016}. They do, however, have
experimental results which shows that the function does not always converge to
a deterministic solution, but it is unclear why the network learns to introduce
more noise into the network parameters.

@@

In this paper we will introduce an extension to NoisyNet that puts it into a Bayesian context. A limitation of NoisyNet is that the initial uncertainty in the Q-value function is crucial to exploration. If the uncertainty is too high, algorithm will struggle to learn anything, while if the uncertainty is too low, there is nothing incentivizing exploration, and the algorithm will likely be stuck in a poor local minimum.

\section{Background}

@@

\begin{itemize}
    \item Optimal regret
    \item Incorporate domain knowledge
    \item Avoid hyperparameter optimization
\end{itemize}

@@

\section{Experiments}
Since our method is essentially an extension of NoisyNet~\cite{fortunato_noisy_2019}, all experiments will compare the performance results of these two methods.
\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare in environments without the advantage of prior knowledge. We hypothesize that the added KL-term will push the algorithm to keep exploring more than NoisyNet, and

\begin{figure}
    \includegraphics[width=\columnwidth]{Example_Figure_1}
    \caption{Example figure 1}
    \label{fig:fig_1}
\end{figure}

\subsection{Informative Prior}
